---
title: "Phan tich Forest plot các ratio của left hand"
author: "Minh"
date: "2024-02-28"
output:
  word_document: default
  html_document:
    keep_md: yes
  pdf_document: default
params:
  warning: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r}
options(warn = -1)
```


## Load bộ số liệu và các packages

```{r}
library(tidyverse)
library(gtsummary)
library(rstatix)
library(readxl)
library(boot)
library(forestplot)
d <- read_excel("C:/Users/Admin/OneDrive/NC_Quangpho/Data_regardless_sleep_new.xlsx")


d$BMI_25_30 <- ifelse(d$BMI_index < 25, 1, 
                         ifelse(d$BMI_index >= 25 & d$BMI_index < 30, 2, 3))
d$Age_stage <- ifelse(d$Age <= 30, 1, 
                         ifelse(d$Age >= 30 & d$Age < 40, 2, ifelse(d$Age >= 40 & d$Age < 50, 3,4)))
d$Gender <- ifelse(d$Gender == 0, "female", ifelse(d$Gender == 1, "male", NA))
d$Race <- d$`6. Race`
d <- d %>%
  mutate(Age_stage = case_when(
    Age_stage == 1 ~ "Age <30",
    Age_stage == 2 ~ "Age [30-40)",
    Age_stage == 3 ~ "Age [40-50)",
    Age_stage == 4 ~ "Age >=50",
    TRUE ~ NA_character_  # This line handles any unexpected values
  ))
d$Hypertension <- ifelse(d$Hypertension == 0, "No", ifelse(d$Hypertension == 1, "Yes", NA))
d$`Stress index` <- ifelse(d$`Stress index` == 0, "No", ifelse(d$`Stress index` == 1, "Yes", NA))
d$Obesity <- ifelse(d$Obesity == 0, "No", ifelse(d$Obesity == 1, "Yes", NA))
d <- d %>%
  mutate(BMI_25_30 = case_when(
    BMI_25_30 == 1 ~ "BMI <25",
    BMI_25_30 == 2 ~ "BMI [25-30)",
    BMI_25_30 == 3 ~ "BMI >=30",
    TRUE ~ NA_character_  # This line handles any unexpected values
  ))
```

## So sánh khác biệt về ratio của chỉ số Stress vs non stress theo các biến

```{r}

# List of variables to iterate over
vars <- c("M", "Kv100", "δ", "T", "A365", "A460", "Anadn", "POM", "A-E", "A-N", "A-M", "A-R", "A-C", "F-E", "F-N", "F-M", "F-R", "F-C")
subgroups <- c("Gender", "Age_stage", "Race", "Type_of_skins", "Stress_index", "Smoking", "Hypertension","BMI_25_30", "Obesity")
```


Hàm để tạo bảng tổng hợp kết quả và p-value

```{r}
compare_ratio_means <- function(data, variable, subgroup, level1, level2, n_bootstrap = 1000) {
  # Filter the data for the two subgroups
  group1_data <- data[data[[subgroup]] == level1, ]
  group2_data <- data[data[[subgroup]] == level2, ]
  
  # Calculate means for the variable of interest in each subgroup
  mean_group1 <- mean(group1_data[[variable]], na.rm = TRUE)
  mean_group2 <- mean(group2_data[[variable]], na.rm = TRUE)
  
  # Calculate the ratio of means
  ratio_means <- mean_group1 / mean_group2
  
  # Perform the Wilcoxon test
  test_result <- wilcox.test(group1_data[[variable]], group2_data[[variable]], paired = FALSE, na.action = na.exclude)
  
  # Bootstrap for CI of the ratio
  bootstrap_ratios <- replicate(n_bootstrap, {
    sample1 <- sample(group1_data[[variable]], replace = TRUE)
    sample2 <- sample(group2_data[[variable]], replace = TRUE)
    mean(sample1, na.rm = TRUE) / mean(sample2, na.rm = TRUE)
  })
  
  CI_lower <- quantile(bootstrap_ratios, probs = 0.025)
  CI_upper <- quantile(bootstrap_ratios, probs = 0.975)
  
  # Construct and return the result as a list
  return(list(
    Variable = variable,
    Subgroup = subgroup,
    Level1_Mean = mean_group1,
    Level2_Mean = mean_group2,
    Ratio = ratio_means,
    CI_Lower = CI_lower,
    CI_Upper = CI_upper,
    P_Value = test_result$p.value
  ))
}


```


```{r}
create_summary_table <- function(data, variables, subgroup, level1, level2, n_bootstrap = 1000) {
  # Initialize an empty list to store the results
  results_list <- list()
  
  # Loop through each variable in the variables list
  for (variable in variables) {
    # Apply the compare_ratio_means function to each variable
    result <- compare_ratio_means(data, variable, subgroup, level1, level2, n_bootstrap)
    # Append the result to the results_list
    results_list[[variable]] <- result
  }
  
  # Combine all the results into a single data frame
  results_df <- do.call(rbind, lapply(results_list, as.data.frame))
  # Optionally, you can convert the row names to a new 'Variable' column if you want
  results_df$Variable <- rownames(results_df)
  # Return the final data frame
  return(results_df)
}

```





Hàm sắp xếp các cột theo thứ tự
```{r}
sort_subgroups <- function(data, characteristic_col, subgroup_col) {
  # Load the dplyr package
  library(dplyr)
  
  # Ensure that the characteristic column is treated as a factor with its levels in the original order
  data[[characteristic_col]] <- factor(data[[characteristic_col]], levels = unique(data[[characteristic_col]]))
  
  # Arrange the subgroups within each characteristic
  sorted_data <- data %>%
    group_by(across(all_of(characteristic_col))) %>%
    arrange(as.numeric(as.character(data[[subgroup_col]])), .by_group = TRUE)
  
  # Return the sorted data
  return(sorted_data)
}


```

Hàm vẽ Forest plot

```{r}
draw_forest_plot <- function(results_table) {
  # Ensure the results_table is not empty
  if(nrow(results_table) == 0) {
    cat("No data available for plotting.\n")
    return()
  }

  # Calculate the number of variables (or rows to be plotted)
  num_vars <- nrow(results_table)

  # Set a base height per variable/row
  base_height_per_var <- 0.5
  
  # Calculate the total plot height, accounting for the additional header row
  total_plot_height <- (num_vars + 1) * base_height_per_var

  # Create the label matrix for the forest plot
  label_matrix <- with(results_table, {
    cbind(
      Variable_Subgroup = paste(Variable),
      Mean_CI = sprintf("%.2f (%.2f-%.2f)", Ratio, CI_Lower, CI_Upper),
      P_Value = sapply(P_Value, function(p) {
        if(p < 0.001) "p < 0.001" else sprintf("p = %.3f", p)
      })
    )
  })

  # Define the header
  header <- rbind(c("Variable", "Mean (95% CI)", "P-value"))

  # Combine header with label matrix
  label_matrix_with_header <- rbind(header, label_matrix)

  # Extract the means and confidence intervals, adding NA for header alignment
  means <- c(NA, results_table$Ratio)
  lower_CI <- c(NA, results_table$CI_Lower)
  upper_CI <- c(NA, results_table$CI_Upper)

  # Generate the forest plot with dynamic height and header
  forestplot(labeltext = label_matrix_with_header, 
             mean = means,
             lower = lower_CI,
             upper = upper_CI,
             xlab = expression("Favour level2 subgroup" * phantom(10) * "Favour level1 subgroup"),
             title = paste("Forest plot of ", results_table$Subgroup, " Ratios in left hand", sep=""),
             zero = 1,  
             line.margin = .1,
             new_page = TRUE,
             col = fpColors(box = "royalblue", line = "darkblue", summary = "royalblue"),
             boxsize = 0.5,
             fig.height = total_plot_height)
}


```

```{r}
variables <- c("M_left", "Kv100_left", "δ_left", "T_left", "A365_left", "A460_left", "Anadn_left", "POM_left", "A-E_left", "A-N_left", "A-M_left", "A-R_left", "A-C_left", "F-E_left", "F-N_left", "F-M_left", "F-R_left", "F-C_left")

```


Bảng Stress index Ratio, level1 = "Yes", level2 = "No"
```{r}
stress_table <- create_summary_table(data = d, variables = variables, subgroup = "Stress index", level1="Yes", level2="No")
stress_table %>% knitr::kable()
```

Forest plot cho Stress index ratio

```{r}
draw_forest_plot(stress_table)
```

Bảng Gender Ratio, level1 = "male", level2 = "female"

```{r}
Gender_table <- create_summary_table(data = d, variables = variables, subgroup = "Gender", level1="male", level2="female")
Gender_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Gender_table)
```


Bảng Smoking ratio, level1 = "Cigarettes, pipe tobacco...", level2 = "Do not smoke"
```{r}
Smoking_table <- create_summary_table(data = d, variables = variables, subgroup = "Smoking", level1="Cigarettes, pipe tobacco...", level2="Do not smoke")
Smoking_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Smoking_table)
```


Bảng Smoking ratio, level1 = "Cigarettes, pipe tobacco...", level2 = "Do not smoke"
```{r}
Smoking_table <- create_summary_table(data = d, variables = variables, subgroup = "Smoking", level1="Cigarettes, pipe tobacco...", level2="Do not smoke")
Smoking_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Smoking_table)
```

Bảng Hypertension ratio, level1 = "Yes", level2 = "No"
```{r}
Hypertension_table <- create_summary_table(data = d, variables = variables, subgroup = "Hypertension", level1="Yes", level2="No")
Hypertension_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Hypertension_table)
```

Bảng Obesity ratio, level1 = "Yes", level2 = "No"
```{r}
Obesity_table <- create_summary_table(data = d, variables = variables, subgroup = "Obesity", level1="Yes", level2="No")
Obesity_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Obesity_table)
```

Bảng Race ratio, level1 = "White", level2 = "Asian or Asian British"
```{r}
Race1_table <- create_summary_table(data = d, variables = variables, subgroup = "Race", level1="White", level2="Asian or Asian British")
Race1_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Race1_table)
```


Bảng Race ratio, level1 = "Black, African, Caribbean or Black Bristish", level2 = "Asian or Asian British"
```{r}
Race2_table <- create_summary_table(data = d, variables = variables, subgroup = "Race", level1="Black, African, Caribbean or Black Bristish", level2="Asian or Asian British")
Race2_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Race2_table)
```

Bảng BMI ratio, level1 = "BMI <25", level2 = "BMI [25-30)"
```{r}
BMI1_table <- create_summary_table(data = d, variables = variables, subgroup = "BMI_25_30", level1="BMI <25", level2="BMI [25-30)")
BMI1_table %>% knitr::kable()
```

```{r}
draw_forest_plot(BMI1_table)
```

Bảng BMI ratio, level1 = "BMI >=30", level2 = "BMI [25-30)"
```{r}
BMI2_table <- create_summary_table(data = d, variables = variables, subgroup = "BMI_25_30", level1="BMI >=30", level2="BMI [25-30)")
BMI2_table %>% knitr::kable()
```

```{r}
draw_forest_plot(BMI2_table)
```

Bảng Age_stage ratio, level1 = "Age <30", level2= "Age [30-40)"

```{r}
Agestage1_table <- create_summary_table(data = d, variables = variables, subgroup = "Age_stage", level1="Age <30", level2="Age [30-40)")
Agestage1_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Agestage1_table)
```

Bảng Age_stage ratio, level1 = "Age <30", level2= "Age [40-50)"

```{r}
Agestage2_table <- create_summary_table(data = d, variables = variables, subgroup = "Age_stage", level1="Age <30", level2="Age [40-50)")
Agestage2_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Agestage2_table)
```

Bảng Age_stage ratio, level1 = "Age <30", level2= "Age >=50"

```{r}
Agestage3_table <- create_summary_table(data = d, variables = variables, subgroup = "Age_stage", level1="Age <30", level2="Age >=50")
Agestage3_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Agestage3_table)
```

Bảng Age_stage ratio, level1 = "Age [30-40)", level2= "Age >=50"

```{r}
Agestage4_table <- create_summary_table(data = d, variables = variables, subgroup = "Age_stage", level1="Age [30-40)", level2="Age >=50")
Agestage4_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Agestage4_table)
```

Bảng Age_stage ratio, level1 = "Age [30-40)", level2= "Age [40-50)"

```{r}
Agestage5_table <- create_summary_table(data = d, variables = variables, subgroup = "Age_stage", level1="Age [30-40)", level2="Age [40-50)")
Agestage5_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Agestage5_table)
```

Bảng Age_stage ratio, level1 = "Age [40-50)", level2= "Age >=50"

```{r}
Agestage6_table <- create_summary_table(data = d, variables = variables, subgroup = "Age_stage", level1="Age [40-50)", level2="Age >=50")
Agestage6_table %>% knitr::kable()
```

```{r}
draw_forest_plot(Agestage6_table)
```
