---
title: "Phan tich Forest plot của right hand/left hand ratio"
author: "Minh"
date: "2024-02-28"
output:
  word_document: default
  html_document:
    keep_md: yes
  pdf_document: default
params:
  warning: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r}
options(warn = -1)
```


## Load bộ số liệu và các packages

```{r}
library(tidyverse)
library(gtsummary)
library(rstatix)
library(readxl)
library(boot)
library(forestplot)
d <- read_excel("C:/Users/Admin/OneDrive/NC_Quangpho/Data_regardless_sleep_18_03.xlsx")
View(d)

d$BMI_25_30 <- ifelse(d$BMI_index < 25, 1, 
                         ifelse(d$BMI_index >= 25 & d$BMI_index < 30, 2, 3))
d$Age_stage <- ifelse(d$Age <= 30, 1, 
                         ifelse(d$Age >= 30 & d$Age < 40, 2, ifelse(d$Age >= 40 & d$Age < 50, 3,4)))
d$Gender <- ifelse(d$Gender == 0, "female", ifelse(d$Gender == 1, "male", NA))
d$Race <- d$`6. Race`
d <- d %>%
  mutate(Age_stage = case_when(
    Age_stage == 1 ~ "Age <30",
    Age_stage == 2 ~ "Age [30-40)",
    Age_stage == 3 ~ "Age [40-50)",
    Age_stage == 4 ~ "Age >=50",
    TRUE ~ NA_character_  # This line handles any unexpected values
  ))
d$Hypertension <- ifelse(d$Hypertension == 0, "No", ifelse(d$Hypertension == 1, "Yes", NA))
d$`Stress index` <- ifelse(d$`Stress index` == 0, "No", ifelse(d$`Stress index` == 1, "Yes", NA))
d$Obesity <- ifelse(d$Obesity == 0, "No", ifelse(d$Obesity == 1, "Yes", NA))
d <- d %>%
  mutate(BMI_25_30 = case_when(
    BMI_25_30 == 1 ~ "BMI <25",
    BMI_25_30 == 2 ~ "BMI [25-30)",
    BMI_25_30 == 3 ~ "BMI >=30",
    TRUE ~ NA_character_  # This line handles any unexpected values
  ))
d$Smoking <- ifelse(d$Smoking == "Do not smoke", "Non smoker", ifelse(d$Smoking == "Cigarettes, pipe tobacco...", "Smoker", NA))
```

## So sánh khác biệt về ratio của chỉ số Right index/ Left index theo các biến

```{r}

# List of variables to iterate over
vars <- c("M", "Kv100", "δ", "T", "A365", "A460", "Anadn", "POM", "A-E", "A-N", "A-M", "A-R", "A-C", "F-E", "F-N", "F-M", "F-R", "F-C")
subgroups <- c("Gender", "Age_stage", "Race", "Typer of skins", "Stress index", "Smoking", "Hypertension","BMI_25_30", "Obesity")
```

Hàm để so sánh lograrit(right/left) với 0 bằng Wilcox test 
```{r}
wilcox_test_func <- function(data, var) {
  ratio <- data[[paste0(var, "_right")]] / data[[paste0(var, "_left")]]
  logtrans <- log(ratio)
  logtrans <- logtrans[!is.infinite(logtrans)]  # Remove infinite values
  
  # Bootstrap function to get the median
  boot_func <- function(data, indices) {
    sample_data <- data[indices]  # Resample with replacement
    return(median(sample_data, na.rm = TRUE))
  }

  # Apply bootstrap
  boot_results <- boot(logtrans, boot_func, R = 1000)  # Increase R for more accuracy
  ci <- boot.ci(boot_results, type = "perc")$percent[4:5]

  # Convert log CI back to ratio
  ci_ratio <- exp(ci)

  test_result <- wilcox.test(logtrans, mu = 0, paired = FALSE, exact = FALSE)
  
  return(data.frame(
    Median_Ratio = median(ratio, na.rm = TRUE),
    CI_Lower = ci_ratio[1],
    CI_Upper = ci_ratio[2],
    P_Value = test_result$p.value
  ))
}
```


Hàm để tạo bảng tổng hợp kết quả và p-value

```{r}
# Function to apply Wilcoxon test across variables for a single subgroup
create_table_for_subgroup <- function(data, vars, subgroup) {
  results_list <- list()
  
  subgroup_levels <- unique(data[[subgroup]])
  
  for(subgroup_level in subgroup_levels) {
    for(var in vars) {
      subgroup_data <- data[data[[subgroup]] == subgroup_level, ]
      
      if(nrow(subgroup_data) > 0) {
        result <- wilcox_test_func(subgroup_data, var)
        result$Subgroup <- subgroup
        result$Subgroup_Level <- as.character(subgroup_level)
        result$Variable <- var
        results_list[[paste(subgroup, subgroup_level, var, sep = "_")]] <- result
      }
    }
  }
  
  results <- do.call(rbind, results_list)
  results <- results[, c("Variable", "Subgroup", "Subgroup_Level", "Median_Ratio", "CI_Lower", "CI_Upper", "P_Value")]
  
  return(results)
}

```

Hàm sắp xếp các cột theo thứ tự
```{r}
sort_subgroups <- function(data, characteristic_col, subgroup_col) {
  # Load the dplyr package
  library(dplyr)
  
  # Ensure that the characteristic column is treated as a factor with its levels in the original order
  data[[characteristic_col]] <- factor(data[[characteristic_col]], levels = unique(data[[characteristic_col]]))
  
  # Arrange the subgroups within each characteristic
  sorted_data <- data %>%
    group_by(across(all_of(characteristic_col))) %>%
    arrange(as.numeric(as.character(data[[subgroup_col]])), .by_group = TRUE)
  
  # Return the sorted data
  return(sorted_data)
}


```

Hàm vẽ Forest plot

```{r}
draw_forest_plot <- function(results_table) {
  # Ensure the results_table is not empty
  if(nrow(results_table) == 0) {
    cat("No data available for plotting.\n")
    return()
  }

  # Calculate the number of variables (or rows to be plotted)
  num_vars <- nrow(results_table)

  # Set a base height per variable/row
  base_height_per_var <- 0.5
  
  # Calculate the total plot height, accounting for the additional header row
  total_plot_height <- (num_vars + 1) * base_height_per_var

  # Create the label matrix for the forest plot
  label_matrix <- with(results_table, {
    cbind(
      Variable_Subgroup = paste(Variable, Subgroup_Level, sep = " - "),
      Mean_CI = sprintf("%.2f (%.2f-%.2f)", Median_Ratio, CI_Lower, CI_Upper),
      P_Value = sapply(P_Value, function(p) {
        if(p < 0.001) "p < 0.001" else sprintf("p = %.3f", p)
      })
    )
  })

  # Define the header
  header <- rbind(c("Variable - Subgroup", "Mean (95% CI)", "P-value"))

  # Combine header with label matrix
  label_matrix_with_header <- rbind(header, label_matrix)

  # Extract the means and confidence intervals, adding NA for header alignment
  means <- c(NA, results_table$Median_Ratio)
  lower_CI <- c(NA, results_table$CI_Lower)
  upper_CI <- c(NA, results_table$CI_Upper)

  # Generate the forest plot with dynamic height and header
  forestplot(labeltext = label_matrix_with_header, 
             mean = means,
             lower = lower_CI,
             upper = upper_CI,
             xlab = expression("Favour left hand" * phantom(10) * "Favour right hand"),
             title = "Forest plot of Right/Left Index Ratios",
             zero = 1,  
             line.margin = .1,
             new_page = TRUE,
             col = fpColors(box = "royalblue", line = "darkblue", summary = "royalblue"),
             boxsize = 0.5,
             fig.height = total_plot_height)
}


```

```{r}
# Define the list of subgroups
subgroup_list <- c("Gender", "Age_stage", "Race", "Typer of skins", "Stress index", "Smoking", "Hypertension","BMI_25_30", "Obesity")
```

Bảng All

```{r}
All_table <- create_table_for_subgroup(d, vars, "All")
All_table<- sort_subgroups(All_table, "Variable", "Subgroup_Level")
knitr::kable(All_table)
```

Các đồ thị Forest plot theo dưới nhóm All

```{r}
for(level in unique(All_table$Subgroup_Level)) {
    level_table <- All_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```



Bảng Gender

```{r}
Gender_table <- create_table_for_subgroup(d, vars, "Gender")
Gender_table<- sort_subgroups(Gender_table, "Variable", "Subgroup_Level")
knitr::kable(Gender_table)
```

Các đồ thị Forest plot theo dưới nhóm Gender

```{r}
for(level in unique(Gender_table$Subgroup_Level)) {
    level_table <- Gender_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```



Bảng Age_stage
```{r}
Age_stage_table <- create_table_for_subgroup(d, vars, "Age_stage")
Age_stage_table<- sort_subgroups(Age_stage_table, "Variable", "Subgroup_Level")
knitr::kable(Age_stage_table)
```

Các đồ thị Forest plot dưới nhóm Age_stage
```{r}
for(level in unique(Age_stage_table$Subgroup_Level)) {
    level_table <- Age_stage_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```


Bảng Race

```{r}
Race_table <- create_table_for_subgroup(d, vars, "Race")
Race_table<- sort_subgroups(Race_table, "Variable", "Subgroup_Level")
knitr::kable(Race_table)
```

Các đồ thị Forest plot theo dưới nhóm Race

```{r}
for(level in unique(Race_table$Subgroup_Level)) {
    level_table <- Race_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```


Bảng Skin

```{r}
Skin_table <- create_table_for_subgroup(d, vars, "Typer of skins")
knitr::kable(Skin_table)
```

Các đồ thị Forest plot theo dưới nhóm Skin

```{r}
for(level in unique(Skin_table$Subgroup_Level)) {
    level_table <- Skin_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```


Bảng Stress

```{r}
Stress_table <- create_table_for_subgroup(d, vars, "Stress index")
Stress_table<- sort_subgroups(Stress_table, "Variable", "Subgroup_Level")
knitr::kable(Stress_table)
```

Các đồ thị Forest plot theo dưới nhóm Stress

```{r}
for(level in unique(Stress_table$Subgroup_Level)) {
    level_table <- Stress_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```


Bảng Smoking

```{r}
Smoking_table <- create_table_for_subgroup(d, vars, "Smoking")
Smoking_table<- sort_subgroups(Smoking_table, "Variable", "Subgroup_Level")
knitr::kable(Smoking_table)
```

Các đồ thị Forest plot theo dưới nhóm Smoking

```{r}
for(level in unique(Smoking_table$Subgroup_Level)) {
    level_table <- Smoking_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```


Bảng Hypertension

```{r}
Hypertension_table <- create_table_for_subgroup(d, vars, "Hypertension")
Hypertension_table<- sort_subgroups(Hypertension_table, "Variable", "Subgroup_Level")
knitr::kable(Hypertension_table)
```

Các đồ thị Forest plot theo dưới nhóm Hypertension

```{r}
for(level in unique(Hypertension_table$Subgroup_Level)) {
    level_table <- Hypertension_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```


Bảng BMI_25_30

```{r}
BMI_25_30_table <- create_table_for_subgroup(d, vars, "BMI_25_30")
BMI_25_30_table<- sort_subgroups(BMI_25_30_table, "Variable", "Subgroup_Level")
knitr::kable(BMI_25_30_table)
```

Các đồ thị Forest plot theo dưới nhóm BMI_25_30

```{r}
for(level in unique(BMI_25_30_table$Subgroup_Level)) {
    level_table <- BMI_25_30_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```


Bảng Obesity

```{r}
Obesity_table <- create_table_for_subgroup(d, vars, "Obesity")
Obesity_table<- sort_subgroups(Obesity_table, "Variable", "Subgroup_Level")
knitr::kable(Obesity_table)
```

Các đồ thị Forest plot theo dưới nhóm Obesity

```{r}
for(level in unique(Obesity_table$Subgroup_Level)) {
    level_table <- Obesity_table %>% dplyr::filter(Subgroup_Level == level)
    
    # Explicitly print the plot
    plot_object <- draw_forest_plot(level_table)
    print(plot_object)}
```



